
---
title: "An Experiment with _'OrthoPanels'_"
author: Davor Cubranic and Mark Pickup
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_width: 7
    fig_height: 5
  rmarkdown::pdf_document:
    toc: false
vignette: >
  %\VignetteIndexEntry{An experiment with _OrthoPanels_}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Let's investigate the accuracy of _opm_'s parameter estimates on 200
simulated datasets.

First, let's define the parameters used by the data-generating
process:
```{r parameters}
rho <- .5
beta <- .5
sig2 <- 1
```

The following function generates a synthetic dataset of desired
dimensions (_N_ cases and _T_ time points) and
distribution parameters ($\rho = \texttt{rho}$, $\beta =
\texttt{beta}$, and $\sigma^2 = \texttt{sig2}$):

```{r generate_fn}
generate <- function(N, T, rho, beta, sig2) {
    f <- runif(N, -2, 2)
    K <- length(beta)
    beta <- matrix(beta, K, 1)

    x <- array(.75*f, dim=c(N, K, T)) + rnorm(N*K*T, sd = 1)

    y <- matrix(0, N, T)
    for (t in 1:T) {
        yy <- if (t>1) y[,t-1] else 0
        y[,t] <- rho * yy + f  + x[,,t] %*% beta + rnorm(N, sd = sqrt(sig2))
    }

    data.frame(i = rep(seq(N), T),
               t = rep(seq(T), each = N),
               as.data.frame(matrix(aperm(x, c(1, 3, 2)), N*T, K,
                                    dimnames = list(NULL, paste0('x', seq(K))))),
               y = c(y))
}
```

Now we generate a dataset with _N_=1000 cases and _T_=3 time points
and fit the model to it 200 times:
```{r simulate}
library(OrthoPanels)
library(knitr)

N <- 1000
T <- 3
reps <- 200

set.seed(123)
opms <- replicate(n = reps,
                  opm(y~x1,
                      data = generate(N = N, T = T,
                                      rho = rho,
                                      beta = beta,
                                      sig2 = sig2),
                      n.samp = 1000),
                  simplify = FALSE)
```

Let's check the sampled parameters:


```{r avg_param}
true_param <- c(rho = rho, sig2 = sig2, beta = beta)
est_param <- sapply(opms, coef)
resid <- sweep(est_param, 1, true_param)
rmse <- sqrt(rowMeans(resid^2))
kable(rbind(`True` = true_param,
            `Est` = rowMeans(est_param),
            `Bias` = rowMeans(resid),
            `RMSE` = rmse))
```

Density plot for each parameter, with true value marked with a
vertical line:

```{r rho_dens}
plot(density(sapply(opms, coef)[1,]),
     main = 'Density of median of posterior samples of rho')
abline(v = rho, col='darkred')
```

```{r sig2_dens}
plot(density(sapply(opms, coef)[2,]),
     main = 'Density of median of posterior samples of sig2')
abline(v = sig2, col='darkred')
```

```{r beta_dens}
plot(density(sapply(opms, coef)[3,]),
     main = 'Density of median of posterior samples of beta')
abline(v = beta, col='darkred')
```

The proportion of time the 95% credible interval includes the true value of the
parameter:

```{r confint}
cis <- sapply(lapply(opms, confint),
             function(ci) {
                 ci[,'2.5%'] <= c(rho, sig2, beta) &
                 ci[,'97.5%'] >= c(rho, sig2, beta)
             })
kable(rowSums(cis) / reps)
```
